---
title: "Clustering"
author: "Lingyi"
date: "6/6/2019"
output: html_document
---
```{r}
knitr::opts_chunk$set(
  cache = TRUE,
  warning = FALSE,
  message = FALSE,
  cache.lazy = FALSE)

dataPath   <- "./GSE120575Data/"
outPath <- "./output/"

library(data.table)
library(dplyr)
library(readxl)
```

```{r}
data <- fread("./GSE120575Data/GSE120575_Sade_Feldman_melanoma_single_cells_TPM_GEO.txt", sep="\t", header = TRUE, fill = TRUE)
data.frame <- as.data.frame(data)
rname <- data.frame[, 1]
dfAll <- data.frame[, -1]
row.names(dfAll) <- rname
dfAll$V16293 <- NULL
```

```{r}
S6 <- read_excel("./GSE120575Data/tableS6.xlsx", sheet="Protein coding genes-methods")
S6.genes <- unlist(S6, use.names=FALSE)
df <- dfAll[S6.genes, ]
```

Find genes with variance > 6
```{r}
# around 40min to run 
index <- which(sapply(S6.genes, function(x) var(as.numeric(df[x,]))>6))
```

```{r}
clu <- df[index, ]
write.csv(clu, file = "./GSE120575Data/clusterData.csv")
```

```{r}
test <- t(clu[1:1000, 1:100])
```


## To determine the optimal number of clusters 
(1) examined how much of the complexity each cluster captures by applying the elbow method 
Select the solutions that are near plateau (k = 10,..., 15)

```{r}
Data <- read.csv(file="./GSE120575Data/clusterData.csv", header=TRUE, sep=",")
genes <- Data[, 1]
rownames(Data) <- genes
Data <- Data[, -1]
```

```{r}
clu <- t(Data)
```

```{r, fig.width=7, fig.height=5}
wss <- 0
# Initialize total within sum of squares error: wss
# For 1 to 15 cluster centers
for (i in 3:15) {
  km.out <- kmeans(clu, centers = i, nstart = 1)
  # Save total within sum of squares to wss variable
  wss[i] <- km.out$tot.withinss
}

# Plot total within sum of squares vs. number of clusters
plot(3:15, wss[3:15], type = "b", 
     xlab = "Number of Clusters", 
     ylab = "Within groups sum of squares")
# Set k equal to the number of clusters corresponding to the elbow location
# k <- 2  # 3 is probably OK, too
```

```{r}
# cluster cells 
# elbow method 
wss <- 0

plot(1:15, unlist(wss, use.names = FALSE),
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```

(2) performed differentical expression anlysis to search for gene markers that are siginificatly more highly expressed in a specific cluster as compared to all other clusters 
(excluded solutions with clusters that have too few marker genes (< 20) distinguishing between them and the rest of the cells. )

```{r}
# FMatrix.RNR <- function(gene){
#   A <- length(which(as.numeric(dfR[gene,]) > 2))
#   B <- length(which(as.numeric(dfNR[gene,])>2))
#   C <- length(which(as.numeric(dfR[gene,]) <= 2))
#   D <- length(which(as.numeric(dfNR[gene,]) <= 2))
#   mean_G1 <- mean(as.numeric(dfR[gene,]))
#   mean_G2 <- mean(as.numeric(dfNR[gene,]))
#   perR <- (A / (A+C))
#   perNR <- (B / (B+D))
#   FMatrix <- matrix(c(A, B, C, D), nrow = 2, dimnames = list(c("G1","G2"), c("log2(TPM+1)>2","log2(TPM+1)<=2")))
#   p <- fisher.test(FMatrix)[["p.value"]]
#   p
#   compute <- log2(mean_G1/mean_G2)
#   out <- data.frame("p_value" = p, "mean_exp_in_R" = mean_G1, "mean_exp_in_NR" = mean_G2, 
#                     "per_in_R" = perR, "per_in_NR" = perNR, "log2_R_NR_" = log2(mean_G1/mean_G2))
#   out
# }
```

(3) Finally, we performed a robustness analysis and selected the clustering solution with the highest median robustness score.

```{r}

```




















